---
title: "Relandscaping Analysis Template"
format: html
editor: visual
---

## Relandscaping Exploratory Analysis: Sully & Thomas & Audrey

## Winter 2024 - MCRS Advanced Quantitative Methods Class Project

Team #:

Team Members: Erin, Sully, Tom, Audrey

Load Packages

```{r}
library(tidyverse)
```

#### Import your Primary file from your saved .Rdata

```{r}
load("RelandscapingDataClean/Size_Primary.Rdata")
```

### Analysis

Before you begin, write your plan for analyzing this data. Add a header for each step you anticipate, then commit and push the document to Git. This may include steps such as:

-   calculating summarized variables (i.e., rates or change between time points, averages of replicated observations)

-   making figures to illuminate comparisons or relationships

-   testing assumptions and/or running models to test for significant comparisons or relationships

Never save over your Primary file. While you are playing around, only make edits under your own workspace header and on your own branch. By doing this, when you commit and push your changes to your shared repository, you avoid merge conflicts.

Once you have something good, show it to your teammates and you can agree to add it (by pasting into the chunk below) to your shared Team Analysis. The team member who updates the Team Analysis should then merge their branch with the master, and all teammates should do a fresh "pull", which will merge the new changes to your current working version. (if this is confusing, ask us for help)

#### Team Analysis

```{r}

```

#### Erin Workspace

```{r}
ggplot(Size_2018_2019, aes(x = as.character(Year), y=Height_cm, color=Treatment, fill = Treatment)) + 
  geom_bar(stat = "identity")+
  labs(title = "Height of Plants Each Year", x = "Year", y = "Height")


ggplot(Size_2018_2019, aes(x = Plant_Species, y=Height_cm, color=Plant_Species, fill = Plant_Species)) + 
  geom_bar(stat = "identity")

```

#### Sully's Workspace

#### How many individual plants are there for each species?

```{r}
# Count up number of individuals in each species
Size_Summary = Size_2018_2019 |> 
  group_by(Plant_Species) |> 
  summarize(count = n())

# Calculate mean number of individuals per species 
Mean_num_individuals = mean(Size_Summary$count)

# Plot 
Species_Count_Plot = Size_Summary |>
  ggplot(aes(x = Plant_Species, y = count, fill = Plant_Species)) + 
  geom_bar(stat = 'identity') + 
  scale_x_discrete(guide = guide_axis(angle = 90)) +
  ylim(0, 38) +
  theme(legend.position = "none") + 
  labs(title = paste(
    'Number of individuals per species (mean: ',
    round(Mean_num_individuals, 2),
    ')',
    sep = '' 
    )) + 
  xlab("Species") + 
  ylab("Count") +
  geom_hline(yintercept = Mean_num_individuals, color="black", linetype="dashed") +
  annotate("text", x = "ELYCON", y = 35, label = "Average", vjust = -0.5, color="maroon") + 
  annotate("segment", x = "ELYCON", xend = "ELYCON", y = Mean_num_individuals, yend = 35, color = "maroon")

Species_Count_Plot

#ggsave(plot = Species_Count_Plot, width = 5, height = 3.5, filename="size-eda-species-count-plot.jpg")


```

#### How big are each of these plants species?

```{r}
# Pivot sizes so that each measurement is its own observation, labeled with height, length or width as is appropriate
Measurements <- Size_2018_2019 |> 
  select(-c(Notes, Date)) |>
  pivot_longer(cols = c("Height_cm", "Width_cm", "Length_cm"), names_to = "Measurement_Type", values_to = "Measurement_cm")

# Create labels for the legend 
Measurement_labels = c("Height", "Length", "Width")
names(Measurement_labels) = c("Height_cm", "Length_cm", "Width_cm")

# Plot, with facets for native / non-native species
Measurements_Plot = Measurements |> 
  ggplot() + aes(x = Plant_Species, y = Measurement_cm, fill = Measurement_Type) + 
  geom_boxplot() + 
  facet_grid(Year ~ Treatment, scales = 'free_x') + 
  scale_x_discrete(guide = guide_axis(angle = 90)) +
  labs(
    title = 'Plant Species Height, Length, and Width',
    fill = 'Measurement Type'
    ) + 
  scale_fill_discrete(labels = Measurement_labels) +
  xlab("Species") + 
  ylab("Measurement (cm)") + 
  theme(legend.position = "top")

Measurements_Plot

#ggsave(plot = Measurements_Plot, width = 8, height = 8, filename="size-eda-measurements-plot.jpg")

```

These graphs are cluttered, but give a good picture of plant sizes across species. I am noticing visually that non-native plants have less variation in their sizes. I can also see which species are taller than they are wide / long, and which are more square. I am curious to learn more about ELAPUN, which is the largest Non-Native plant, and whether its size has an impact on the ecosystem.

### In-class EDA

```{r}

Size_2018_2019 = Size_2018_2019 |> mutate(
  Volume = Height_cm * Length_cm * Width_cm,
  Year = factor(Year)
)

Volume_Plot = Size_2018_2019 |> 
  ggplot(aes(x = Treatment, y = log(Volume), color = Year)) +
  geom_boxplot() + 
  labs(
     title = 'Year-over-year Plant Volume'
   ) + 
   xlab("Plant Species")

Volume_Plot

#ggsave(plot = Volume_Plot, width = 7, height = 5, filename="plant-volume.jpg")

  
```

```{r}

Size_Morphological_Pairs = Size_2018_2019 |> mutate(
  Morphological_Pair = case_when(
    Plant_Species == "ROSOFF" 
    | Plant_Species == "ERIFAS" 
    ~ "ROSOFF-ERIFAS", 
    Plant_Species == "RHAIND" 
    | Plant_Species == "FRACAL" 
    ~ "RHAIND-FRACAL",
    Plant_Species == "LAVGIN" 
    | Plant_Species == "SALAPI" 
    ~ "LAVGIN-SALAPI",
    Plant_Species == "ATLFES" 
    | Plant_Species == "PENSPA"
    | Plant_Species == "MUHRIG"
    | Plant_Species == "ELYCON" 
    ~ "GRASS",
    Plant_Species == "ACHTOM" 
    | Plant_Species == "ACHMIL" 
    ~ "ACHTOM-ACHMIL",
    .default = NA
  )
) |> filter(!is.na(Morphological_Pair))

Morphological_Plot = Size_Morphological_Pairs |> group_by(Morphological_Pair) |> 
   ggplot(aes(x = Plant_Species, y = log(Volume), color = Year, fill = Treatment)) +
   geom_boxplot() + 
   scale_fill_manual(values = c("gray", "white")) +
   facet_wrap( 
     ~ Morphological_Pair, scales = 'free_x',
     labeller = as_labeller(c(
       'GRASS' = 'ELYCON and PENSPA (grasses)',
       'ROSOFF-ERIFAS' = 'ERIFAS and ROSOFF',
       'LAVGIN-SALAPI' = 'LAVGIN and SALAPI',
       'RHAIND-FRACAL' = 'RHAIND and FRACAL'
     ))
     ) + 
   labs(
     title = 'Year-over-year Plant Volume for Morphological Pairs'
   ) + 
   xlab("Plant Species")

Morphological_Plot

#ggsave(plot = Morphological_Plot, width = 7, height = 5, filename="plant-volume-morphological-pairs.jpg")
   
```

```{r}

Pair_Plot_Height = Size_Morphological_Pairs |> group_by(Morphological_Pair) |> 
   ggplot(aes(x = Plant_Species, y = Height_cm, color = Year, fill = Treatment)) +
   geom_boxplot() + 
   scale_fill_manual(values = c("gray", "white")) +
   facet_wrap( 
     ~ Morphological_Pair, scales = 'free_x',
     labeller = as_labeller(c(
       'GRASS' = 'ELYCON and PENSPA (grasses)',
       'ROSOFF-ERIFAS' = 'ERIFAS and ROSOFF',
       'LAVGIN-SALAPI' = 'LAVGIN and SALAPI',
       'RHAIND-FRACAL' = 'RHAIND and FRACAL'
     ))
     ) + 
   labs(
     title = 'Year-over-year Plant Height for Morphological Pairs'
   ) + 
   xlab("Plant Species")

Pair_Plot_Height

#ggsave(plot = Pair_Plot_Height, width = 7, height = 5, filename="plant-height-morphological-pairs.jpg")
```

```{r}
Pair_Plot_Length = Size_Morphological_Pairs |> group_by(Morphological_Pair) |> 
   ggplot(aes(x = Plant_Species, y = Length_cm, color = Year, fill = Treatment)) +
   geom_boxplot() + 
   scale_fill_manual(values = c("gray", "white")) +
   facet_wrap( 
     ~ Morphological_Pair, scales = 'free_x',
     labeller = as_labeller(c(
       'GRASS' = 'ELYCON and PENSPA (grasses)',
       'ROSOFF-ERIFAS' = 'ERIFAS and ROSOFF',
       'LAVGIN-SALAPI' = 'LAVGIN and SALAPI',
       'RHAIND-FRACAL' = 'RHAIND and FRACAL'
     ))
     ) + 
   labs(
     title = 'Year-over-year Plant Length for Morphological Pairs'
   ) + 
   xlab("Plant Species")

Pair_Plot_Length

#ggsave(plot = Pair_Plot_Length, width = 7, height = 5, filename="plant-length-morphological-pairs.jpg")
```

```{r}
Pair_Plot_Width = Size_Morphological_Pairs |> group_by(Morphological_Pair) |> 
   ggplot(aes(x = Plant_Species, y = Width_cm, color = Year, fill = Treatment)) +
   geom_boxplot() + 
   scale_fill_manual(values = c("gray", "white")) +
   facet_wrap( 
     ~ Morphological_Pair, scales = 'free_x',
     labeller = as_labeller(c(
       'GRASS' = 'ELYCON and PENSPA (grasses)',
       'ROSOFF-ERIFAS' = 'ERIFAS and ROSOFF',
       'LAVGIN-SALAPI' = 'LAVGIN and SALAPI',
       'RHAIND-FRACAL' = 'RHAIND and FRACAL'
     ))
     ) + 
   labs(
     title = 'Year-over-year Plant Width for Morphological Pairs'
   ) + 
   xlab("Plant Species")

Pair_Plot_Width

#ggsave(plot = Pair_Plot_Height, width = 7, height = 5, filename="plant-width-morphological-pairs.jpg")
```
